---
title: "Presentation Exercise"
author: "Holly Milazzo"
---



In this exercise, I recreated the bar graph that illustrates the share of political donations from team owners in six major sports leagues (NFL, NBA, WNBA, NHL, MLB, and NASCAR) to the Republican and Democratic parties over three election years: 2016, 2018, and 2020 from the FiveThirtyEight page here [Link](https://fivethirtyeight.com/features/inside-the-political-donation-history-of-wealthy-sports-owners/). The graph shows a significant majority of donations going to the Republican party in all three years. The annotation highlights that Giants owner Charles Johnson's contributions constitute a substantial portion of the total Republican donations, underscoring the influence of individual donors in political contributions.


```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(dplyr)
library(ggplot2)
library(readr)
library(tidyr)
```


```{r}
# Important
# Define the data location
data_location_sports <- here::here("presentation-exercise", "sports-political-donations.csv")

# Read the CSV data
rawdata_sports <- read_csv(data_location_sports, col_names = FALSE)

# Preview the data
head(rawdata_sports)
```


```{r, message=FALSE}

# Assign proper column names
colnames(rawdata_sports) <- c("Owner", "Team", "League", "Recipient", "Amount", "Election_Year", "Party")

# Clean up the Amount column (remove '$' and convert to numeric)
rawdata_sports$Amount <- as.numeric(gsub("[\\$,]", "", rawdata_sports$Amount))

# Filter out rows where Party is not 'Democrat' or 'Republican'
rawdata_sports <- rawdata_sports[rawdata_sports$Party %in% c("Democrat", "Republican"), ]

# Summarize data to calculate total donations by Party and Year
donations_summary <- aggregate(rawdata_sports$Amount, 
                               by = list(Election_Year = rawdata_sports$Election_Year, Party = rawdata_sports$Party), 
                               FUN = sum)
colnames(donations_summary)[3] <- "Total_Amount"

# Convert Party to factor and specify order (Democrat first for better visualization)
donations_summary$Party <- factor(donations_summary$Party, levels = c("Democrat", "Republican"))

# Calculate the proportion of total donations by year
total_by_year <- aggregate(donations_summary$Total_Amount, 
                           by = list(Election_Year = donations_summary$Election_Year), 
                           FUN = sum)
colnames(total_by_year)[2] <- "Yearly_Total"

donations_summary <- merge(donations_summary, total_by_year, by = "Election_Year")
donations_summary$Prop_Amount <- donations_summary$Total_Amount / donations_summary$Yearly_Total

# Plot the data
ggplot(donations_summary, aes(x = as.factor(Election_Year), y = Prop_Amount, fill = Party)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("Democrat" = "blue", "Republican" = "red")) +
  labs(
    title = "Team owners give largely to the GOP",
    subtitle = "Share of donations from team owners in six leagues, per year and party",
    x = "",
    y = "Share of Donations",
    fill = "Donations to"
  ) +
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "gray90"),
    plot.background = element_rect(fill = "gray90"),
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  ) +
  annotate("text", x = 1.5, y = 1.05, label = "Giants owner Charles Johnson’s total contributions make up 32.1% of all Republican contributions.", size = 3, hjust = 0.3, vjust = 0)


```

Table: Partisan Contributions by League

The table breaks down the total political contributions from owners and commissioners in six sports leagues (MLB, NBA, NHL, NFL, WNBA, and NASCAR) to the Republican and Democratic parties from 2016 to 2020. It reveals that MLB owners have donated the most overall, with significant amounts going to both parties, but predominantly to Republicans. The table provides a clear comparison of partisan contributions across different leagues, highlighting the disparity in political support within the sports industry.

Both visuals emphasize the substantial financial support sports team owners provide to political parties, predominantly favoring the Republican party, and the considerable influence of a few key donors.

```{r}
# Load necessary libraries
library(ggplot2)
library(ggtext)

# Define the data for the table
data_table <- data.frame(
  League = c("MLB", "NBA", "NHL", "NFL", "WNBA", "NASCAR"),
  To_Republicans = c("$15,181,761", "$8,372,300", "$7,087,116", "$5,032,470", "$1,338,459", "$576,110"),
  To_Democrats = c("$5,184,604", "$2,641,487", "$1,726,733", "$873,500", "$1,634,153", "$93,983"),
  Total = c("$20,366,365", "$11,013,787", "$8,813,849", "$5,905,970", "$2,972,612", "$670,093")
)

# Create a base plot
p <- ggplot(data_table, aes(x = 1, y = League)) +
  geom_text(aes(label = To_Republicans, x = 2), hjust = 0, color = "red") +
  geom_text(aes(label = To_Democrats, x = 3), hjust = 0, color = "blue") +
  geom_text(aes(label = Total, x = 4), hjust = 0) +
  geom_text(aes(label = League, x = 0), hjust = 0, fontface = "bold") +
  scale_x_continuous(limits = c(-0.5, 4.5), breaks = 0:4, labels = c("LEAGUE", "TO REPUBLICANS", "TO DEMOCRATS", "TOTAL", "")) +
  theme_void() +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 10, hjust = 0.5),
    plot.caption = element_text(size = 8, hjust = 0.5),
    axis.text.x = element_text(size = 10, face = "bold"),
    plot.background = element_rect(fill = "white", color = "white"),
    panel.background = element_rect(fill = "white", color = "white")
  ) +
  labs(
    title = "MLB owners have donated the most",
    subtitle = "Specifically partisan contributions from owners and commissioners in the NFL, NBA, WNBA, NHL, MLB and NASCAR, by party, 2016-20",
    caption = "SOURCE: FEDERAL ELECTION COMMISSION, OPENSECRETS"
  )

# Print the plot
print(p)


```




